%\documentclass[conference]{IEEEtran}
%\documentclass{sig-alternate}
\documentclass{sig-alternate-05-2015}

\usepackage{datatool}
%\usepackage{amsthm}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{enumerate}
\usepackage{cprotect}
\usepackage{xtab}
\usepackage{tikz}
\usepackage{url}
\usepackage{hyperref}
\usepackage{cite}
\usepackage[update,prepend,outdir=./]{epstopdf}
\DTLsetseparator{•}
\DTLloaddb{data}{../analysis_output/database.csv}

\usepackage{balance}

%submission site: http://cyberchairpro.borbala.net/asepapers/submit/
% Login: carl1978@iastate.edu
% Password: ASE-143111797041267

\widowpenalty=10000
\clubpenalty=10000

\newcommand{\todoNow}[1]{\textbf{\textcolor{red}{TODO.NOW: #1}}} %comment out for submission
\newcommand{\todoMid}[1]{\textbf{\textcolor{magenta}{TODO.MID: #1}}} %comment out for submission
%\newcommand{\todoMid}[1]{\textbf{\textcolor{magenta}{}}} %comment out for submission
\newcommand{\todoLast}[1]{\textbf{\textcolor{blue}{TODO.LAST: #1}}} %comment out for submission
\newcommand{\yes}{\tikz\draw[black,fill=black] (0,0) circle (.5ex);}
\newcommand{\sorta}{\tikz\draw[gray,fill=gray] (0,0) circle (.5ex);}
\newcommand{\no}{\tikz\draw[black] (0,0) circle (.5ex);}
\renewcommand{\dbltopfraction}{.9}
\usepackage{bigstrut}
    \setlength\bigstrutjot{2pt}
\newtheorem{mydef}{Definition}

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
%\title{Regex Feature Use In Practice}
%\title{Regular Expression Feature Usage in Python}
\title{Exploring  Regular Expression Usage and Context\\ in Python}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\numberofauthors{2}

%\author{
%\alignauthor
%Carl Chapman\\
%       \affaddr{Department of Computer Science}\\
%       \affaddr{Iowa State University}\\
%       \affaddr{Ames, IA, USA}\\
%      \email{carl1978@iastate.edu}\\
% \alignauthor
%Kathryn T. Stolee\\
%       \affaddr{Departments of Computer Science}\\
%       \affaddr{North Carolina State University}\\
%       \affaddr{Raleigh, NC, USA}\\
%      \email{ktstolee@ncsu.edu}\\
%}

% make the title area
\maketitle


\begin{abstract}
%Regular expressions are used frequently in programming languages for form validation, ad-hoc file searches, and simple parsing.
Due to the popularity and pervasive use of regular expressions, researchers have created tools to support their creation, validation, and use. However, little is known about the context in which regular expressions are used, the features that are most common, and how behaviorally similar regular expressions are to one another.
%This information is critical to inform tool designers about how to best support developers working with regular expressions.
%Each tool has made design decisions about which regular expression features to support, and these decisions impact the usefulness and power of the tools. Yet, these decisions are often made with little information as there does not exist an empirical study of regular expression feature usage to inform these design decisions.

In this paper, we explore the context in which regular expressions are used through a combination of developer surveys and repository analysis.
We survey 18 professional developers from a small software company about their regular expression usage and pain points.
Our results indicate that developers frequently use regular expressions in their programming practices, often composing regular expressions at least weekly.
Then, we
analyzed nearly 4,000 open source Python projects from GitHub and extracted nearly 14,000 unique regular expression patterns, focusing the analysis on how often features are used.
We  map the most common features used in regular expressions to those features supported by four common regular expression engines from industry and academia: brics, Hampi, RE2, and Rex.
Using similarity analysis of regular expressions across projects,
we identify six common behavioral clusters that describe how regular expressions are often used in practice.
This is the first rigorous examination of regex usage and it provides empirical evidence to support design decisions by regex tool
builders. It also points to areas of needed future work, such as refactoring regular expressions to increase refactoring understandability, support for migrating regexes between language, and context-specific tool support for common regexes usages.
%We conclude by discussing the implications for  tool  designers and outline several directions of future work.
%\todoLast{Katie: revise}



%Due to the popularity and pervasive use of regular expressions, researchers have created tools to support their creation, validation, and use. However, little is known about the context in which regular expressions are used, the features that are most common, and how behaviorally similar regular expressions are to one another. This information is critical to inform tool designers about how to best support developers working with regular expressions.
%
%In this paper, we explore the context in which regular expressions are used. We survey 18 professional developers about the context and frequency of their regular expression usage. Then, we explore a sample of regular expressions, focusing on how often distinct features are used.  We analyzed nearly 4,000 open source Python projects from GitHub and extracted nearly 14,000 unique regular expression patterns that were used for analysis. We also map the most common features used in regular expressions to those features supported by four common regular expression engines from industry and academia: brics, Hampi, RE2, and Rex. Our results indicate that developers frequently use regular expressions in their programming practices, often composing regular expressions at least weekly. Using semantic analysis of regular expressions across projects, we identify six common behavioral clusters that describe how regular expressions are often used in practice. We conclude by discussing the implications for  tool  designers and outline several directions of future work.

%FYI: nProjScanned = 3898
%These observations lead to suggestions on which regular expression features are imperative to support and for future directions in research to help developers write and reuse regular expressions.
\end{abstract}

\input{introduction}

\input{related}

\input{survey}

\input{study}

\input{results}

\input{discussion}

\section{Threats to Validity}
\label{sec:threats}
The following threats impact our results and conclusions:

%\subsection{Conclusion}
\textbf{Reliability of Measures:} The validity of our survey results is dependent on the clarity of the questions. The authors went through several iterations of the survey and included examples for all the regex feature descriptions to improve understandability.

The similarity measure between regexes used in the cluster algorithm is computed empirically rather than analytically, and the more Rex-generated strings used to compute the similarity measure, the more likely it is to be accurate. Our experiments used 400 strings to balance performance and precision, but a higher number could lead to more cohesive clusters. Additionally, regex patters that use any feature not supported by Rex were omitted from the cluster analysis. Last, the threshold of 0.75 was chosen based on the MCL recommendation, but it may not create optimal clusters. %\todoMid{talk about tools for formal analysis}
%In our clustering algorithm, we require a similarity level of 0.75 so that we do not over-estimate the similarity between regular  expressions.\todoNow{what does this last sentence mean?  Is this paragraph okay?}

%\subsection{Internal}

\textbf{Instrumentation:} Regular expression patterns were clustered using strings generated by the Rex tool.  We assume that the strings generated by Rex are reasonably diverse to help characterize the regex behavior. To mitigate this threat, Rex generated 400 strings per regex and we inspected strings randomly to ensure diversity.

Implementation errors are a risk for research involving repository analysis. To combat this, we have tested our code and made the repository publicly available\footnote{\url{https://github.com/softwarekitty/tour_de_source}}.

\textbf{Selection:} We mined \DTLfetch{data}{key}{nProjScanned}{value} Python projects from GitHub, which is  small compared to all available Python projects. The projects were mined using the GitHub API which sorts the projects by creation date. By using the API, the goal was to reduce any sampling bias introduced by the researchers.

We also did not scrape all commits in every project for regular expression utilizations, rather, we grabbed each project every 20 commits. It is possible that in between the scanned commits, a regex utilization was added and then removed, leading to fewer utilizations in our final data set.

\textbf{Recall bias:} Survey participants were asked to reflect on their past behavior, which may not  represent actual behavior. To mitigate this, we designed the survey to lead participants to think about their behavior before summarizing (e.g., asking how often they use regexes in each of several environments before asking about usage frequency).

%\subsection{Construct}

\textbf{Hypothesis guessing:} The lead author was an employee of Dwolla when the survey was run, so participant behavior may have been affected by their relationship with the author. To combat this, the beginning of the survey stated, ``As this is for research, please answer as accurately as possible, not guessing what is the desired answer and providing that."

%\subsection{External}

\textbf{Interaction of Selection and Treatment:} Our survey participants were software developers from a small startup company and may not be representative of developers who use regexes. Given that the average participant has nine years of development experience, their responses likely pull from a variety of experiences with regular expression usage.

\textbf{Interaction of Setting and Treatment:}
We only explore regular expressions in Python projects so these results may be coupled with the activities performed using Python and not generalize to other languages. The regex usage context reported by survey participants, however, includes information on  regex usage in a  variety of settings and languages. Future work will replicate this study in other languages and the survey with a larger group of participants.



%The survey results are from a set of developers at one company and may not generalize to other developers at that company or at other companies. While the results provide some information on regex usage, repeating the survey to a larger set of diverse developers will increase the generalizeability of the results.
%%
%
%In extracting patterns for analysis, we omit utilizations that contain flags since flags provide refinements on the functionality of the pattern as it is used in the project. Given that only 12.7\% of all utilizations used flags, the impact on the clusters should be minimal.
%
%%The feature analysis is dependent on the features identified in the PCRE parser.
%


\vspace{-9pt}
\section{Conclusion}
\label{sec:conclusion}
In this work, we have explored the contexts in which regular expressions are used as well as the features and behavioral similarities of regexes found in open source Python projects. In a survey of 18 professional developers, we find that 50\% compose regular expressions at least weekly. The most common purposes are locating content within a file or capturing parts of strings. The most difficult parts about working with regular expressions were reported to be composing and reading them.
In a study of regular expression usage in nearly 4,000 Python projects, we find that over 42\% of projects contain a regular expression.
We present an approach to measure behavioral similarity between regexes by generating strings that match one regex and pairwise  testing the remaining regexes against it. This similarity measure is used to form cross-project behavioral clusters. In the top 100 largest clusters, we find that capturing the contents of brackets,  searching for delimiter characters and matching alternative values were common behaviors.
These results have implications for tool designers and for future research aimed at better supporting developers in using regular expressions.


% use section* for acknowledgement
\vspace{-9pt}
\section*{Acknowledgment}
Special thanks to the Dwolla developers for their survey participation.
This work is supported in part by NSF SHF-1218265, NSF SHF-EAGER-1446932, and the Harpole-Pentair endowment at Iowa State University.

\vspace{12pt}

\bibliographystyle{abbrv}

% The bibliography should be embedded for final submission.
\balance
\bibliography{biblio}



% that's all folks
\end{document}

